### 2019-11-30 writing/proposal.tex
Rename directories and start flask app

\documentclass{article}

\title{WorldGAN: 3D Environment Generation Using Deep Adversarial Networks}

\author{Tim Whitaker}

\begin{document}

	\maketitle

	\section{Introduction}

	Generative adversarial networks are a class of machine learning models that have become very popular since their inception in 2014. They've been the engine behind some of the most popular demos in the field, including Deep Fakes, This Person Does Not Exist, and Text-to-Image translation. These models are powerful in generating new and creative content, and I think they could be a good aid to humans in designing 3 dimensional environments for video games.

	I am proposing a method of allowing a user to draw a top-down 2-dimensional segmentation map, using a GUI akin to MS Paint and using a generative adversarial network to translate that into a 3-dimensional environment in Unreal Engine.

	\section{Background}

	The motivation for this project is to help game designers build environments and worlds easier and more quickly. This method acts as a form of pseudo-procedural generation, and would allow for an endless amount of variation in content, without needing a user to manually define these variations.

	Generative adversarial networks have only been applied to 3 dimensions very recently, and this work has only been done with modeling 3d objects like chairs and tables. The idea for this research is mostly inspired by the Nvidia GauGAN project http://nvidia-research-mingyuliu.com/gaugan/ and the relevant topics I'll be researching include procedural generation and conditional image synthesis.

	\subsection{Procedural Generation}

	Procedural generation has been used in the gaming industry for decades. It allows game designers to introduce variation to a game without needing to explicitly build each variant by hand. This can introduce millions of options on both the environment and the content, increasing the replayability value of a game and leading to a novel experience on each playthrough. This method is very popular in dungeon crawlers like nethack and rogue, and more recently it's been applied to massive 3-dimensional worlds like No Man's Sky and Minecraft.

	\subsection{Conditional Image Synthesis}

	Conditional image synthesis is the process of generating an image conditionally from a dataset based on a class label. The two big categories are text-to-image and image-to-image translation. The famous problems of style transfer, black and white photo colorization, and super resolution all fall into this category. GANs are a class of models that are often used for the task of conditional image synthesis.

	\subsection{Literature Review}

	The following is a collection of literature that are relevant to this project. Most of the content here deals with GANs and procedural generation.

	\begin{itemize}
		\item Physically Based Rendering: From Theory to Implementation\\
		https://www.pbrt.org/
		\item Semantic Image Synthesis with Spatially-Adaptive Normalization\\
		https://arxiv.org/abs/1903.07291
		\item Procedural Content Generation via Machine Learning (PCGML)\\
		https://arxiv.org/abs/1702.00539	
		\item Generative Adversarial Networks\\
		https://arxiv.org/abs/1406.2661
		\item Unsupervised Representation Learning with Deep Convolutional Generative Adversarial Networks\\
		https://arxiv.org/abs/1511.06434
		\item Conditional Generative Adversarial Networks\\
		https://arxiv.org/abs/1411.1784
		\item Image-to-Image Translation with Conditional Adversarial Networks\\
		https://arxiv.org/abs/1611.07004		
	\end{itemize}

	\section{Methodology}

	I see this project broken up into three parts. First is the user interface. A simple GUI akin to MS Paint is sufficient and will allow a user to draw a 2d top-down semantic map. Second is creating and training the GAN to generate the terrain heightmap/mesh from the user input. Third is generating all of the style, textures, and content, like trees and rocks, that will be placed in/on the terrain. Due to the nature of Unreal Engine, I'm not sure if this step can be done with a machine learning model. I will need to experiment with Unreal Engine more to see if it's possible and if it's not, then more traditional forms of procedural generation may be used for this step.

	I will need to build up a set of training data to bootstrap my GAN. I will do this manually by drawing both the input and output heightmaps using photoshop. It may be possible to augment this process by using volunteers and/or generating the inputs programmatically and creating the outputs manually.

	\subsection{Deliverables}

	\begin{itemize}
		\item A simple user interface to allow a user to draw a 2d map using semantically significant brushes.
		\item A dataset of 2d images and heightmaps to train the GAN on.
		\item A GAN that will translate that 2d map to a 3d heightmap/mesh.
		\item A method for applying texture and objects to the 3d heightmap in Unreal Engine.
	\end{itemize}
\end{document}

### 2019-11-30 app/backend/server.py

@@ -1,0 +1,7 @@
+from flask import Flask
+
+app = Flask(__name__)
+
+@app.route("/")
+def hello():
+  return "Hello world"

### 2019-12-04 app/backend/server.py
Serve ui app through flask

@@ -1,7 +1,0 @@
-from flask import Flask
-
-app = Flask(__name__)
-
-@app.route("/")
-def hello():
-  return "Hello world"

### 2019-12-04 app/server.py

@@ -1,0 +1,7 @@
+from flask import Flask, render_template
+
+app = Flask(__name__)
+
+@app.route("/")
+def hello():
+  return render_template("index.html")

### 2019-12-04 utilities/data-augmentator/main.py
Create augmentations for original batch

@@ -1,0 +1,31 @@
+import os
+from PIL import Image, ImageOps
+
+in_directory = "../../data/working/usermaps"
+out_directory = "../../data/augmented-batch/"
+
+def augment(in_directory, out_directory):
+  for file in os.listdir(in_directory):
+    if file.endswith(".png"):
+      path = os.path.join(in_directory, file)
+      im = Image.open(path)
+
+      im.save(out_directory + file)
+
+      flipped = ImageOps.flip(im)
+      flipped.save(out_directory + "f_" + file)
+
+      mirror = ImageOps.mirror(im)
+      mirror.save(out_directory + "m_" + file)
+
+      ninety = im.rotate(90)
+      ninety.save(out_directory + "90_" + file)
+
+      oneeighty = im.rotate(180)
+      oneeighty.save(out_directory + "180_" + file)
+
+      twoseventy = im.rotate(270)
+      twoseventy.save(out_directory + "270_" + file)
+
+augment("../../data/original-batch/cycle/trainA", "../../data/augmented-batch/trainA/")
+augment("../../data/original-batch/cycle/trainB", "../../data/augmented-batch/trainB/")

### 2019-12-05 writing/paper/paper.tex
Start paper

\documentclass[twocolumn]{article}

\title{WorldGAN: 3D Environment Generation Using Deep Adversarial Networks}

\author{Tim Whitaker}

\begin{document}

	\maketitle

	\section{Abstract}

	\section{Introduction}

	\subsection{Objectives}

	\section{Background}

	\section{Methodology}

	Building the UI app.

	Building the training set.

	First round of training for pix2pix and cyclegan.

	Augmenting the training set.

	Second round of training for pix2pix and cyclegan.

	Building the test set.

	Visualizations of results with 3d software.

	\section{Results}

	\section{Conclusion}

\end{document}

### 2019-12-06 writing/paper/paper.tex
Work on the paper

@@
 
-\author{Tim Whitaker}
+\author{Tim Whitaker \\ Colorado State University}
 
@@
 
+	We introduce a method and application for generating 3D environments with generative adversarial networks.
+The outputs from our networks are able to be imported into popular 3d modeling software like unreal, unity or blender.
+We've created a simple and easy to use web based interface to allow a user to control the output of the generated environment by painting using predefined brushes that represent terrain varieties, mountains, hills, terraces, plains, and rivers.
+
 	\section{Introduction}
 
-	\subsection{Objectives}
+	This project explores the usage of conditional generative adversarial networks to synthesize 3 dimensional environments.
+
+	The motivations for this project are to make it easier for game, world or environment designers to create realistic environments procedurally while giving the designer more control over the results.
+
+	We aim to create a simple and easy to use GUI interface for a user to paint a 2 dimensional semantically significant image map.
+We then pass the user created image map through a generative adversarial network trained on elevation data gathered from the Earth in order to generate a realistic terrain heightmap that approximately represents the user defined map.
+The resulting heightmap produced from the neural network can then be uploaded into any of the popular 3 dimensional rendering software including Unity, Unreal, Blender or World Machine.
 
@@
 
+	This project involves work touching on a number of fields of computer science. Combining machine learning, procedural generation, and three dimensional content is a relatively new area of research.
+
+	\subsection{Procedural Content Generation}
+
+	Procedural content generation is an algorithmic and stochastic process used in art, games, music and film.
+As a paradigm for content generation, it is used to increase replayability, variability, and flexibility.
+It can reduce production cost and effort and it can save on storage space.
+
+	\subsection{Generative Adversarial Networks}
+
+	Neural networks have become a pervasive part of computer science.
+It allows for a powerful model of learning patterns from complex datasets.
+Generative adversarial networks work by synthesizing novel data according to patterns it learns in the training set.
+We evaluate two models of conditional generative adversarial networks to achieve image to image translation.
+
+	\subsection{Related Work}
+
+	Nvidia GauGAN.
+
 	\section{Methodology}
 
-	Building the UI app.
+	In this section we go over all of the pieces of work that went in to this project.
 
-	Building the training set.
+	\subsection{User Interface Application}
 
-	First round of training for pix2pix and cyclegan.
+	The user interface is a web based javascript application. The application allows a user to paint a semantically significant image using predefined brushes.
 
-	Augmenting the training set.
+	The backend is powered by a Flask API.
 
-	Second round of training for pix2pix and cyclegan.
+	\subsection{The Dataset}
 
-	Building the test set.
+	\subsubsection{Training Data}
+
+	I used an open source map tool https://github.com/tangrams/heightmapper.
+This tool is an interactive application elevation data gathered from Mapzen's global elevation service https://www.mapzen.com/blog/elevation/.
+
+	I then built a program that controls a headless browser using an open source library called puppeteer https://github.com/puppeteer/puppeteer.
+This script loads a random coordinate, checks if we are over land and saves an image of the elevation data at that coordinate.
+We resize the saved image to be 512 x 512 pixels.
+
+	I then started manually building the usermap training set using the user interface application I built.
+For each image in the heightmap training set, I would paint a usermap to represent that heightmap, using the semantically significant brushes that most closely matched the terrain (mountains, hills, plains, rivers, terrace).
+
+	After this process we are left with 258 paired usermaps and heightmaps. We group all of these together into batch one.
+
+	We then created a second batch by taking batch one and augmenting the images to increase the size of a training set by 5x.
+We performed 5 operations on each usermap and heightmap.
+We mirror horizontally, flip vertically, rotate 90 derees, rotate 180 degrees, and rotate 270 degrees.
+
+	Pix2pix requires the training data from both the heightmaps and usermaps need to be paired.
+We stack each pair of training images horizontally and save this new image into our training batch.
+The resulting images are 1024 x 512 pixels.
+
+	\subsubsection{Test Data}
+
+	The test data is a collection of 5 usermaps I drew that are designed to be representativ of a variety of environments a user might draw.
+These 5 environments include a mountainous region, a mountainous region fading into a hilly region, a hilly region fading into a flat region, a region with a main river and many sub rivers running through mountains, and a combined region containing mountains, hills, plains and a river.
+
+	\subsection{Training The Models}
+
+	We trained two models on this initial dataset, Pix2Pix and CycleGAN. GTX 1080ti GPU. i7 8700k. processor.
+
+	\subsubsection{Pix2Pix}
+	~1 hour training time.
+	~4 hours training time.
+
+	\subsubsection{CycleGAN}
+	~4 hours training time.
+	~24 hours training time.
+
+	\subsection{Visualizations Using 3D Software}
 
@@
 
+	\subsection{Visualizations Using 3D Software}
+
 	\section{Conclusion}
 
+	\subsection{Future Work}
+
+	\bibliography{}
+
 \end{document}

### 2019-12-06 writing/paper/paper.tex
Start bib

@@
 	We introduce a method and application for generating 3D environments with generative adversarial networks.
-The outputs from our networks are able to be imported into popular 3d modeling software like unreal, unity or blender.
-We've created a simple and easy to use web based interface to allow a user to control the output of the generated environment by painting using predefined brushes that represent terrain varieties, mountains, hills, terraces, plains, and rivers.
+The outputs from our networks are able to be imported into popular three dimensional modeling software like Unreal, Unity and Blender.
+We've created a simple and easy to use web based interface to allow a user to control the output of the generated environment by painting with predefined brushes that represent terrain varieties: mountains, hills, terraces, plains, and rivers.
 
@@
 We then pass the user created image map through a generative adversarial network trained on elevation data gathered from the Earth in order to generate a realistic terrain heightmap that approximately represents the user defined map.
-The resulting heightmap produced from the neural network can then be uploaded into any of the popular 3 dimensional rendering software including Unity, Unreal, Blender or World Machine.
+The resulting heightmap produced from the neural network can then be uploaded into any of the popular 3 dimensional rendering softwares including Unity, Unreal, Blender or World Machine.
 
@@
 
-	This project involves work touching on a number of fields of computer science. Combining machine learning, procedural generation, and three dimensional content is a relatively new area of research.
+	This project explores the new paradigm of 3 dimensional procedurally generated content via machine learning.
 
@@
 	Procedural content generation is an algorithmic and stochastic process used in art, games, music and film.
-As a paradigm for content generation, it is used to increase replayability, variability, and flexibility.
-It can reduce production cost and effort and it can save on storage space.
+As a paradigm for content generation, it is used to increase replayability, variability, flexibility and reduce production cost, effort storage space \cite{summerville2017procedural}.
 
@@
 We evaluate two models of conditional generative adversarial networks to achieve image to image translation.
+
+	We are going to evaluate two implementations of a specialized form of generative adversarial networks designed to perform conditional image to image translation.
+The two models we are using are called Pix2Pix and CycleGAN.
+
+	\subsubsection{Pix2Pix}
+
+	\subsubsection{CycleGAN}
 
@@
 	\subsection{The Dataset}
+
+	Generative adversarial networks need a good amoount of training data to output good results and a key objective of this project is to generate realistic results.
+This leads to an interesting challenge of putting together a realistic yet large enough dataset.
+In order to achieve this, we use terrain data of the earth as opposed to fictional or manufactured terrain data.
+In the next two sections we go into the methodology behind building the dataset.
 
@@
 
-	\bibliography{}
+	\bibliographystyle{plain}
+	\bibliography{references}
 

### 2019-12-06 writing/paper/paper.tex
Write more sections of the paper

@@
 \documentclass[twocolumn]{article}
+\usepackage{graphicx}
 
@@
 
+	Both Pix2Pix and CycleGAN are based on the deep convolutional generative adversarial network architecture proposd by Radford et al. in 2015 \cite{radford2015unsupervised}.
+
 	\subsubsection{Pix2Pix}
@@
 
-	The user interface is a web based javascript application. The application allows a user to paint a semantically significant image using predefined brushes.
+	The user interface is a web based javascript application that allows a user to paint a semantically significant image using predefined brushes.
+We include tools to change th brush size and export or clear the image.
+The canvas size is 512 x 512 pixels, and these images are resized automatically when passed to our neural network.
 
-	The backend is powered by a Flask API.
+	\begin{figure}[h!]
+	\includegraphics[width=1\columnwidth]{ui.png} 
+	\caption{The user interace application}
+	\end{figure}
+
+	The backend is powered by a Flask API. We set up an endpoint to be able to evaluate a drawn image on our trained models.
 
@@
 	Generative adversarial networks need a good amoount of training data to output good results and a key objective of this project is to generate realistic results.
-This leads to an interesting challenge of putting together a realistic yet large enough dataset.
+This leads to an interesting challenge of putting together a realistic yet sizable dataset.
 In order to achieve this, we use terrain data of the earth as opposed to fictional or manufactured terrain data.
@@
 
-	I used an open source map tool https://github.com/tangrams/heightmapper.
-This tool is an interactive application elevation data gathered from Mapzen's global elevation service https://www.mapzen.com/blog/elevation/.
+	I used an interactive open source map tool that visualizes elevation data gathered from Mapzen's global elevation service \cite{richardson2016mapzen}.
 
-	I then built a program that controls a headless browser using an open source library called puppeteer https://github.com/puppeteer/puppeteer.
-This script loads a random coordinate, checks if we are over land and saves an image of the elevation data at that coordinate.
+	I then built a program that controls a headless browser using an open source library developed by Google called puppeteer.
+This script loads a random coordinate, checks if the coordinate is over land and if so, then we save an image of the elevation data at that coordinate.
 We resize the saved image to be 512 x 512 pixels.
 
-	I then started manually building the usermap training set using the user interface application I built.
-For each image in the heightmap training set, I would paint a usermap to represent that heightmap, using the semantically significant brushes that most closely matched the terrain (mountains, hills, plains, rivers, terrace).
+	We then started manually building the usermap training set using the WorldGAN user interface application.
+For each image in the heightmap training set, We paint a usermap to represent that heightmap, using the semantically significant brushes that most closely match the terrain (mountains, hills, plains, rivers, and terraces).
+This process is a subjective one and all usermaps were painted by myself.
+This does bias the model to produce images according to how I read the heightmaps.
 
-	After this process we are left with 258 paired usermaps and heightmaps. We group all of these together into batch one.
+	After this process we are left with 259 paired usermaps and heightmaps. We group all of these together into a batch to be trained on.
 
-	We then created a second batch by taking batch one and augmenting the images to increase the size of a training set by 5x.
+	Creating the usermaps by hand is quite time consuming, and I wanted to experiment with models trained on a large dataset.
+We accomplish this by augmenting our first batch with simple transformations in order to artificially increase the number of images we can train on.
+Image augmentations are a well researched and effective appoach to improving neural network performance \cite{perez2017effectiveness}.
+
 	We performed 5 operations on each usermap and heightmap.
-We mirror horizontally, flip vertically, rotate 90 derees, rotate 180 degrees, and rotate 270 degrees.
+We mirror horizontally, flip vertically, rotate 90 derees, rotate 180 degrees, and rotate 270 degrees.	All of the resulting images are gathered into a second batch and we end up with 1548 paired images (a 5x increase).
 
-	Pix2pix requires the training data from both the heightmaps and usermaps need to be paired.
-We stack each pair of training images horizontally and save this new image into our training batch.
-The resulting images are 1024 x 512 pixels.
+	Before training the networks, we need to perform one more preprocessing step.
+Pix2Pix requires paired images to be combined into a single side by side image.
+We stack each pair of usermap and heightmap and save these images into separate batches spcifically for Pix2Pix.
+Each resulting image is 1024 x 512 pixels.
+CycleGAN does not need this preprocessing step and can be trained on images in different directories.
 
@@
 
-	The test data is a collection of 5 usermaps I drew that are designed to be representativ of a variety of environments a user might draw.
-These 5 environments include a mountainous region, a mountainous region fading into a hilly region, a hilly region fading into a flat region, a region with a main river and many sub rivers running through mountains, and a combined region containing mountains, hills, plains and a river.
+	The test data is a collection of 10 usermaps I drew that are designed to be representative of a variety of environments a user might draw.
+These 5 environments include mountainous regions, hilly regions, flat regions and combined regions.
 
@@
 
-	We trained two models on this initial dataset, Pix2Pix and CycleGAN. GTX 1080ti GPU. i7 8700k. processor.
+	We trained a total of 4 models.
+Pix2Pix and CycleGAN on both the original batch of training data and the augmented batch.
+The training was performed on my desktop GPU enabled PC.
+The relevant specifications include an Nvidia GTX 1080ti GPU with 11 GB of memory, an Intel i7 8700k processor, a 500GB solid state hard drive, and 32 GB of RAM.
 

### 2019-12-06 writing/paper/paper.tex
Write

@@
 
-\title{WorldGAN: 3D Environment Generation Using Deep Adversarial Networks}
+\title{Conditional 3D Environment Generation Using Deep Adversarial Networks}
 
@@
 
+	Pix2Pix introduces a a generalized approach to image to image translation problems.
+Their model learns not only a mapping between input and output images, but it also learns a loss function between the two image datasets.
+This allows this model to be used for a variety of image tasks that would originally need very different model and loss configurations \cite{isola2016imagetoimage}.
+
 	\subsubsection{CycleGAN}
+
+	CycleGAN introduces the idea of cycle consistency loss.
+It works by measuring the loss of a mapping from one distribution to another and then back to the original distribution again.
+This approach allows for good generalization and the advantage of not needing specifically paired images in the training distributions.
+This can save a lot of time in building large datasets, but it can lose out on the specificity offered by Pix2Pix \cite{zhu2017unpaired}.
 
@@
 
+	Both models are built on top of the PyTorch framework \cite{paszke2019pytorch}.
+
 	\subsubsection{Pix2Pix}

### 2019-12-07 app/server.py
Send image to flask server

@@ -1,7 +1,22 @@
-from flask import Flask, render_template
+import base64
+from flask import Flask, request, render_template
+from subprocess import call
 
 app = Flask(__name__)
 
 @app.route("/")
 def hello():
   return render_template("index.html")
+
+@app.route("/generate", methods=['POST'])
+def generate_heightmap():  
+  data = request.get_json()
+  b64 = data['b64']
+
+  with open("generate/a.png", "wb") as fh:
+    fh.write(base64.decodebytes(b64.encode()))
+
+  #call(["ls", "../pytorch-CycleGAN-and-pix2pix"])
+  #call(["python", "../pytorch-CycleGAN-and-pix2pix/test.py"])
+
+  return "hi"

### 2019-12-07 app/static/index.js

@@ -98,3 +98,16 @@ function exportImage() {
 
   w.document.write("<img src='" + data + "' />")
 }
+
+function generate() {
+  const data = canvas.toDataURL("image/png");
+  const image = data.replace("data:image/png;base64,", "");
+
+  var json = JSON.stringify({'b64': image});
+
+  var xhr = new XMLHttpRequest();
+  xhr.open('POST', '/generate', true);
+  xhr.setRequestHeader('Content-Type', 'application/json');
+
+  xhr.send(json);
+}

### 2019-12-07 app/server.py
Send image to flask

@@ -1,5 +1,5 @@
 import base64
-from flask import Flask, request, render_template
+from flask import Flask, request, render_template, send_file
 from subprocess import call
 
 app = Flask(__name__)
@@ -16,7 +16,6 @@ def generate_heightmap():
   with open("generate/a.png", "wb") as fh:
     fh.write(base64.decodebytes(b64.encode()))
 
-  #call(["ls", "../pytorch-CycleGAN-and-pix2pix"])
   #call(["python", "../pytorch-CycleGAN-and-pix2pix/test.py"])
 
-  return "hi"
+  #return send_file("generate/results/b.png", mimetype="image/png")

### 2019-12-07 app/static/index.js

@@ -109,5 +109,15 @@ function generate() {
   xhr.open('POST', '/generate', true);
   xhr.setRequestHeader('Content-Type', 'application/json');
 
+  xhr.onreadystatechange = function() {
+    if (xhr.readyState == 4) {
+      if(xhr.status == 200) {
+        var obj = JSON.parse(xhr.responseText);
+        
+        console.log(obj);
+      }
+    }
+  };
+
   xhr.send(json);
 }

### 2019-12-07 app/server.py
Write code to call model from flask

@@ -13,9 +13,14 @@ def generate_heightmap():
   data = request.get_json()
   b64 = data['b64']
 
-  with open("generate/a.png", "wb") as fh:
+  with open("generate/test/a.png", "wb") as fh:
     fh.write(base64.decodebytes(b64.encode()))
 
-  #call(["python", "../pytorch-CycleGAN-and-pix2pix/test.py"])
+  call(["python3", "../pytorch-CycleGAN-and-pix2pix/test.py", 
+    "--dataroot", "./generate", 
+    "--checkpoints_dir", "../models", 
+    "--name", "two-pix2pix", 
+    "--results_dir", "./generate/results", 
+    "--model", "pix2pix"])
 
   #return send_file("generate/results/b.png", mimetype="image/png")

### 2019-12-07 app/server.py
Finish the app!

@@ -13,7 +13,7 @@ def generate_heightmap():
   data = request.get_json()
   b64 = data['b64']
 
-  with open("generate/test/a.png", "wb") as fh:
+  with open("generate/test/A.png", "wb") as fh:
     fh.write(base64.decodebytes(b64.encode()))
 
   call(["python3", "../pytorch-CycleGAN-and-pix2pix/test.py", 
@@ -23,4 +23,7 @@ def generate_heightmap():
     "--results_dir", "./generate/results", 
     "--model", "pix2pix"])
 
-  #return send_file("generate/results/b.png", mimetype="image/png")
+  with open("generate/results/two-pix2pix/test_latest/images/A_fake_B.png", "rb") as fh:
+      encoded_string = base64.b64encode(fh.read())
+
+  return encoded_string

### 2019-12-07 app/static/index.js

@@ -100,24 +100,28 @@ function exportImage() {
 }
 
 function generate() {
-  const data = canvas.toDataURL("image/png");
-  const image = data.replace("data:image/png;base64,", "");
+  const data = canvas.toDataURL("image/png")
+  const image = data.replace("data:image/png;base64,", "")
 
-  var json = JSON.stringify({'b64': image});
+  var json = JSON.stringify({ b64: image })
 
-  var xhr = new XMLHttpRequest();
-  xhr.open('POST', '/generate', true);
-  xhr.setRequestHeader('Content-Type', 'application/json');
+  var xhr = new XMLHttpRequest()
+  xhr.open("POST", "/generate", true)
+  xhr.setRequestHeader("Content-Type", "application/json")
 
   xhr.onreadystatechange = function() {
     if (xhr.readyState == 4) {
-      if(xhr.status == 200) {
-        var obj = JSON.parse(xhr.responseText);
-        
-        console.log(obj);
+      if (xhr.status == 200) {
+        results = document.getElementById("results")
+        results.innerHTML =
+          '<img src="data:image/png;base64,' + xhr.responseText + '"/>'
       }
     }
-  };
+  }
 
-  xhr.send(json);
+  xhr.send(json)
+
+  //spinner
+  results = document.getElementById("results")
+  results.innerHTML = '<img src="/static/spinner.svg"/>'
 }

### 2019-12-08 writing/paper/paper.tex
Update screenshot and write more on the model sections

@@
 \documentclass[twocolumn]{article}
+
 \usepackage{graphicx}
+\usepackage{amsmath}
+\usepackage{amsfonts}
 
-\title{Conditional 3D Environment Generation Using Deep Adversarial Networks}
+\title{3D Environment Generation Using Conditional Adversarial Networks}
 
@@
 
-	We introduce a method and application for generating 3D environments with generative adversarial networks.
+	We introduce a method and application for generating 3D environments using conditional generative adversarial networks.
 The outputs from our networks are able to be imported into popular three dimensional modeling software like Unreal, Unity and Blender.
-We've created a simple and easy to use web based interface to allow a user to control the output of the generated environment by painting with predefined brushes that represent terrain varieties: mountains, hills, terraces, plains, and rivers.
+We've created a simple web based interface to allow a user to control the output of the generated environment by painting with predefined brushes that represent terrain varieties: mountains, hills, terraces, plains, and rivers.
 
@@
 
-	Neural networks have become a pervasive part of computer science.
-It allows for a powerful model of learning patterns from complex datasets.
+	Neural networks allow for a powerful model of learning patterns from complex datasets.
 Generative adversarial networks work by synthesizing novel data according to patterns it learns in the training set.
@@
 
+	The objective function of a conditional GAN can be expressed as:
+
+	\begin{align*}
+		\mathcal{L}_{cGAN}(G, D) &= \mathbb{E}_{x,y}[log(D(x, y))] \\
+		&+ \mathbb{E}_{x,z}[log(1 - D(x, G(x, z))]
+	\end{align*}
+
+	Where $G$ is the generator network that is attempting to minimize this objective and D is the discriminator network that tries to maximize it.
+
 	\subsubsection{Pix2Pix}
@@
 
+	Pix2Pix found it beneficial to add a traditional loss term as well, namely L1 distance.
+The discriminators job is the same as a traditional GAN, but the generator now is tasked with fooling the discriminator as well as producing results near the ground truth output.
+
+	$$
+		\mathcal{L}_{L1}(G) = \mathbb{E}_{x,y,z}[|| y - G(x, z) ||_1]		
+	$$
+
+	The addition of that loss term can be tuned with a parameter  $\lambda$. This combined objective function is used for training Pix2Pix.
+
+	$$	
+		G^* = arg \mathop{min}_G \mathop{max}_D \mathcal{L}_{cGAN}(G, D) + \lambda \mathcal{L}_{L1}(G)
+	$$
+
 	\subsubsection{CycleGAN}
 
-	CycleGAN introduces the idea of cycle consistency loss.
+	CycleGAN introduces cycle consistency loss to conditional generativeadversarial networks.
 It works by measuring the loss of a mapping from one distribution to another and then back to the original distribution again.
@@
 This can save a lot of time in building large datasets, but it can lose out on the specificity offered by Pix2Pix \cite{zhu2017unpaired}.
+
+	CycleGAN introduces a cycle consistency loss to their objective function.
+The idea is that for each image in the $x$ domain, the image translation cycle should be able to bring the $x$ back to the original image, $x \rightarrow G(x) \rightarrow F(G(x)) \approx x$.
+
+	\begin{align*}
+		\mathcal{L}_{cyc}(G, F) =\ & \mathbb{E}_{x,y}[||F(G(x)) - x||_1]\ + \\
+		& \mathbb{E}_{x,z}[||G(F(y)) - y||_1]
+	\end{align*}
+
+	\begin{align*}
+		\mathcal{L}(G, F, D_X, D_Y) &=\ \mathcal{L}_{GAN}(G, D_Y, X, Y) \\
+		&+\ \mathcal{L}_{GAN}(F, D_X, Y, X) \\
+		&+\ \lambda \mathcal{L}_cyc (G, F)  
+	\end{align*}
 
@@
 	The user interface is a web based javascript application that allows a user to paint a semantically significant image using predefined brushes.
-We include tools to change th brush size and export or clear the image.
+We include tools to change the brush size and export or clear the image.
 The canvas size is 512 x 512 pixels, and these images are resized automatically when passed to our neural network.
@@
 
-	The backend is powered by a Flask API. We set up an endpoint to be able to evaluate a drawn image on our trained models.
+	The backend is powered by a Flask API.
+We set up an endpoint to be able to evaluate a drawn image on our trained models.
+In order to pass the usermap and generated heightmap back and forth between the client and the server, we use base 64 encoding and decoding on the image.
+This way we can save the images to the file system and have a local process run the neural network and return the generated heightmap.
+This direct access to the file system has security concerns, so it's only recommended to run the app locally.
 
@@
 
-	Generative adversarial networks need a good amoount of training data to output good results and a key objective of this project is to generate realistic results.
+	Generative adversarial networks need a good amount of training data to output good results and a key objective of this project is to generate realistic results.
 This leads to an interesting challenge of putting together a realistic yet sizable dataset.
@@
 
-	I used an interactive open source map tool that visualizes elevation data gathered from Mapzen's global elevation service \cite{richardson2016mapzen}.
+	We use an interactive open source map tool that visualizes elevation data gathered from Mapzen's global elevation service \cite{richardson2016mapzen}.
 
-	I then built a program that controls a headless browser using an open source library developed by Google called puppeteer.
+	We then built a program that controls a headless browser using an open source library developed by Google called puppeteer.
 This script loads a random coordinate, checks if the coordinate is over land and if so, then we save an image of the elevation data at that coordinate.
@@
 
-	We then started manually building the usermap training set using the WorldGAN user interface application.
+	We then manually build the usermap training set using our user interface application.
 For each image in the heightmap training set, We paint a usermap to represent that heightmap, using the semantically significant brushes that most closely match the terrain (mountains, hills, plains, rivers, and terraces).
@@
 
-	After this process we are left with 259 paired usermaps and heightmaps. We group all of these together into a batch to be trained on.
+	After this process we are left with 258 paired usermaps and heightmaps. We group all of these together into a batch to be trained on.
 
@@
 	The test data is a collection of 10 usermaps I drew that are designed to be representative of a variety of environments a user might draw.
-These 5 environments include mountainous regions, hilly regions, flat regions and combined regions.
+These 10 environments include mountainous regions, hilly regions, flat regions and combined regions.
 
@@
 	We trained a total of 4 models.
-Pix2Pix and CycleGAN on both the original batch of training data and the augmented batch.
+Pix2Pix and CycleGAN on both the original and augmented batch of training data.
 The training was performed on my desktop GPU enabled PC.
-The relevant specifications include an Nvidia GTX 1080ti GPU with 11 GB of memory, an Intel i7 8700k processor, a 500GB solid state hard drive, and 32 GB of RAM.
+The operating system is Ubuntu 18.04.
+We train on a Nvidia GTX 1080ti GPU with 11 GB of memory, an Intel i7 8700k processor, a 500GB solid state hard drive, and 32 GB of RAM.
 
-	Both models are built on top of the PyTorch framework \cite{paszke2019pytorch}.
+	Both models are built on top of the PyTorch framework \cite{paszke2019pytorch} and are trained for 200 epochs.
 
 	\subsubsection{Pix2Pix}
-	~1 hour training time.
-	~4 hours training time.
+		Our pix2pix model contained a total of 57.183 million parameters.
+The generator network contains 54.414 million parameters and the discriminator contains 2.769 million parameters.
+The original batch of data took approximately 45 minutes to train.
+The augmented batch took approximately 4 hours to train.
 
 	\subsubsection{CycleGAN}
-	~4 hours training time.
-	~24 hours training time.
+		Our cyclegan model contained a total of 28.286 million parameters.
+Generator A and Generator B each contain 11.378 million parameters.
+Discriminator A and Discriminator B each contain 2.765 million parameters.
+The original batch of data took approximately 4 hours to train.
+The augmented batch took approximately 24 hours to train.
 
@@
 
-	Visualizations of results with 3d software.
+	The resulting heightmaps can be imported into popular 3d modeling software including Unity, Unreal, Blender, and World Machine.
+For the purposes of visualizing these results, and because of the limited support for Linux in all of the previously listed softwares, we visualize results using a demo application called Fractscape.
+Fractscape includes elevation based texture mapping which happens to work very well for visualizing our terrain elevation data.
+Unity, Unreal, Blender and World Machine have a different workflow for painting textures, so using the generality of fractscape works best for our visualizations.
 
@@
 
-	\subsection{Visualizations Using 3D Software}
+	Subjective results of networks show that pix2pix performs much better than cyclegan.
+For the specfic task of segmented image translation, paired training data allows the network to learn the meaning of the segmentations much better.
+Cyclegan had issues of keeping the semantic meaning in tact, even though it does appear to generate some of the patterns correctly, it's often mapped to the wrong brush or user defined section.
+
+	Both models trained on the augmented batch produce much higher fidelity images. Both pix2pix models produce usable and realistic terrain.
+
+	We also find that the model does not perform very well on usermaps with rivers.
+We found that rivers tend to produce noisy results and sometimes the rivers aren't defined enough to be noticeable in the results.
+I think this can be solved with more training data including rivers.
+Due to the nature of how the training data was gathered, it's likely that we didn't get enough examples, and due to the subjective nature of myself reading the earth data and drawing the usermaps myself, I could have read them wrong and not painted them as detailed as thy needed to be.
+
+	Terraces also had a lack of training data.
+Due to the specificity of their appearance, they were able to produce better results than rivers.
+The models will likely struggle on terraces with odd shapes.
+
+	Mountains, hills, and plains all performed very well.
+There was plenty of samples in the training for these features and the areas for each tend to be large.
+This allowed the networks to do a great job of modeling these terrain features.
 

### 2019-12-08 writing/paper/paper.tex
Add results table

@@
 \usepackage{amsfonts}
+\usepackage{floatpag}
 
@@
 	Both models are built on top of the PyTorch framework \cite{paszke2019pytorch} and are trained for 200 epochs.
+We use the open source repo implemented by the Jun-Yan Zhu, who published the research on both models \cite{junyanzhu2017github}.
 
@@
 Fractscape includes elevation based texture mapping which happens to work very well for visualizing our terrain elevation data.
-Unity, Unreal, Blender and World Machine have a different workflow for painting textures, so using the generality of fractscape works best for our visualizations.
+Unity, Unreal, Blender and World Machine have a different workflow for painting textures, so using the generality of fractscape works best for our visualizations \cite{starscenesoftware2006}.
 
@@
 For the specfic task of segmented image translation, paired training data allows the network to learn the meaning of the segmentations much better.
-Cyclegan had issues of keeping the semantic meaning in tact, even though it does appear to generate some of the patterns correctly, it's often mapped to the wrong brush or user defined section.
+Cyclegan had issues of keeping the semantic meaning in tact.
+Even though it does appear to generate some of the patterns correctly, it's often mapped to the wrong brush or user defined section.
 
@@
 
+	\begin{table*}[t]
+	\thisfloatpagestyle{empty}
+	\vspace{-2cm}
+		\begin{tabular}{c|c|c|c|c}
+ 			Test Image & Original Pix2Pix & Augmented Pix2Pix & Original CycleGAN & Augmented CycleGAN \\
+ 			\hline
+ 				\includegraphics[trim=0 0 0 -20, width=0.25\columnwidth]{00.png} & \includegraphics[width=0.25\columnwidth]{01.png} & \includegraphics[width=0.25\columnwidth]{02.png} & \includegraphics[width=0.25\columnwidth]{03.png} & \includegraphics[width=0.25\columnwidth]{04.png} \\
+ 				\includegraphics[width=0.25\columnwidth]{10.png} & \includegraphics[width=0.25\columnwidth]{11.png} & \includegraphics[width=0.25\columnwidth]{12.png} & \includegraphics[width=0.25\columnwidth]{13.png} & \includegraphics[width=0.25\columnwidth]{14.png} \\
+ 				\includegraphics[width=0.25\columnwidth]{20.png} & \includegraphics[width=0.25\columnwidth]{21.png} & \includegraphics[width=0.25\columnwidth]{22.png} & \includegraphics[width=0.25\columnwidth]{23.png} & \includegraphics[width=0.25\columnwidth]{24.png} \\
+ 				\includegraphics[width=0.25\columnwidth]{30.png} & \includegraphics[width=0.25\columnwidth]{31.png} & \includegraphics[width=0.25\columnwidth]{32.png} & \includegraphics[width=0.25\columnwidth]{33.png} & \includegraphics[width=0.25\columnwidth]{34.png} \\
+ 				\includegraphics[width=0.25\columnwidth]{40.png} & \includegraphics[width=0.25\columnwidth]{41.png} & \includegraphics[width=0.25\columnwidth]{42.png} & \includegraphics[width=0.25\columnwidth]{43.png} & \includegraphics[width=0.25\columnwidth]{44.png} \\
+ 				\includegraphics[width=0.25\columnwidth]{50.png} & \includegraphics[width=0.25\columnwidth]{51.png} & \includegraphics[width=0.25\columnwidth]{52.png} & \includegraphics[width=0.25\columnwidth]{53.png} & \includegraphics[width=0.25\columnwidth]{54.png} \\
+ 				\includegraphics[width=0.25\columnwidth]{60.png} & \includegraphics[width=0.25\columnwidth]{61.png} & \includegraphics[width=0.25\columnwidth]{62.png} & \includegraphics[width=0.25\columnwidth]{63.png} & \includegraphics[width=0.25\columnwidth]{64.png} \\
+ 				\includegraphics[width=0.25\columnwidth]{70.png} & \includegraphics[width=0.25\columnwidth]{71.png} & \includegraphics[width=0.25\columnwidth]{72.png} & \includegraphics[width=0.25\columnwidth]{73.png} & \includegraphics[width=0.25\columnwidth]{74.png} \\
+ 				\includegraphics[width=0.25\columnwidth]{80.png} & \includegraphics[width=0.25\columnwidth]{81.png} & \includegraphics[width=0.25\columnwidth]{82.png} & \includegraphics[width=0.25\columnwidth]{83.png} & \includegraphics[width=0.25\columnwidth]{84.png} \\
+ 				\includegraphics[width=0.25\columnwidth]{90.png} & \includegraphics[width=0.25\columnwidth]{91.png} & \includegraphics[width=0.25\columnwidth]{92.png} & \includegraphics[width=0.25\columnwidth]{93.png} & \includegraphics[width=0.25\columnwidth]{94.png} \\
+
+		\end{tabular}
+	\end{table*}
+
 	\section{Conclusion}

### 2019-12-10 app/server.py
Rerun p2p test and adjust parameters

@@ -21,7 +21,8 @@ def generate_heightmap():
     "--checkpoints_dir", "../models", 
     "--name", "two-pix2pix", 
     "--results_dir", "./generate/results", 
-    "--model", "pix2pix"])
+    "--model", "pix2pix",
+    "--load_size", "256"])
 
   with open("generate/results/two-pix2pix/test_latest/images/A_fake_B.png", "rb") as fh:
       encoded_string = base64.b64encode(fh.read())

### 2019-12-10 writing/paper/paper.tex
Add related and future work

@@
 
+	This approach offers a promising avenue to co-creative design.
+Humans working together with machine learning programs could allow for more creative results with reduced cost and effort for the designer.
+The flexibility of being able to tweak and tune the training set also allows for the possibility of adapting this approach to content of many different kinds, not just 3d environments.
+
 	\section{Background}
-
-	This project explores the new paradigm of 3 dimensional procedurally generated content via machine learning.
 
@@
 	Procedural content generation is an algorithmic and stochastic process used in art, games, music and film.
-As a paradigm for content generation, it is used to increase replayability, variability, flexibility and reduce production cost, effort storage space \cite{summerville2017procedural}.
+As a paradigm for content generation, it is used to increase replayability, variability, flexibility and reduce production cost, effort and storage space \cite{summerville2017procedural}.
+These algorithms are crafted by hand and they require a lot of tuning by a developer to ensure that the random perturbations result in usable content.
+
+	Procedral content generation has been used in production applications since 1980 with the dungeon crawler Rogue demonstrating its effectiveness.
+\cite{wiki:procedural} Procedural generation has been a staple in gaming since and today it continues to be featured in some of the top games.
+
+	The addition of machine learning to procedural generation is a new topic in research.
+Most projects are using 2 dimensional platformers, but real time strategy and 3 dimensional environments are also being explored.
+This field of research has its own challenges, particularly the comparability of results, but efforts are beng made to improve benchmarks and datasets \cite{summerville2017procedural}.
 
@@
 Generative adversarial networks work by synthesizing novel data according to patterns it learns in the training set.
-We evaluate two models of conditional generative adversarial networks to achieve image to image translation.
+They work by training two neural networks in competition with each other, a generator and a discriminator.
+The generator network generates new data instances and the discriminator classifies the results of the generator according to how authentic the generated images are given some training distribution.
+These two networks train in tandem, and over time, the generator gets good at producing novel images that look like they belong to the training set.
 
-	We are going to evaluate two implementations of a specialized form of generative adversarial networks designed to perform conditional image to image translation.
+	We are evaluate two implementations of a specialized form of generative adversarial networks designed to perform conditional image to image translation.
 The two models we are using are called Pix2Pix and CycleGAN.
 
-	Both Pix2Pix and CycleGAN are based on the deep convolutional generative adversarial network architecture proposd by Radford et al. in 2015 \cite{radford2015unsupervised}.
+	Both Pix2Pix and CycleGAN are conditonal networks based on the deep convolutional generative adversarial network architecture proposed by Radford et al. in 2015 \cite{radford2015unsupervised}.
 
-	The objective function of a conditional GAN can be expressed as:
+	The objective function of a conditional generative adversarial networks can be expressed as:
 
@@
 
-	CycleGAN introduces a cycle consistency loss to their objective function.
+	Where pix2pix trains one GAN (one generator and one discriminator), CycleGAN trains two, (Generator A and B and Discriminator A and B).
+learning the mapping between the two GANs is the trick to getting unpaired image translation to work.
+CycleGAN introduces a cycle consistency loss.
 The idea is that for each image in the $x$ domain, the image translation cycle should be able to bring the $x$ back to the original image, $x \rightarrow G(x) \rightarrow F(G(x)) \approx x$.
@@
 
-	Nvidia GauGAN.
+	\subsubsection{Nvidia's GauGAN}
+
+	GauGAN was the inspiration behind the semantic user map.
+This research demonstrated an application that took a 2 dimensional semantic image drawn by a user, and translated it into a photorealistic image.
+This project also used Pix2Pix as a base for their network architecture, but they also introduce a novel normalization layer called spatially-adaptive normalization.
+The idea is that normal layer normalization tends to wash out semantic meaning from an image.
+By considering the spatial properties, you can keep the semantic information in tact while still performing normalization.
+\cite{park2019semantic}	.
+Nvidia hosts an excellent demo of the project at http://nvidia-research-mingyuliu.com/gaugan.
+
+	\subsubsection{A Step Toward Procedural Terrain Generation with GANs}
+
+	This paper had a very similar idea to this project.
+They use a standard generative adversarial network to synthesize terrain heightmaps with a model trained on Earth data.
+They use NASA's visible earth project for their training image data at a resolution of 1km per pixel.
+This project appears to be quite introductory and they lay out a good overview of how a standard GAN could be implemented to generate environments.
+We expand on their work by introducing the conditional GAN and integrating co-creative design by including a translation from a user defined segmentation map.
+\cite{beckham2017step}
 
 	\section{Methodology}
-
-	In this section we go over all of the pieces of work that went in to this project.
 
@@
 I think this can be solved with more training data including rivers.
-Due to the nature of how the training data was gathered, it's likely that we didn't get enough examples, and due to the subjective nature of myself reading the earth data and drawing the usermaps myself, I could have read them wrong and not painted them as detailed as thy needed to be.
+Due to the nature of how the training data was gathered, it's likely that we didn't get enough examples, and due to the subjective nature of myself reading the earth data and drawing the usermaps myself, I could have read them wrong and not painted them as detailed as they needed to be.
 
@@
 This allowed the networks to do a great job of modeling these terrain features.
+
+	\subsection{User Preference Survey}
+
+	We also created a small user survey to measure user preference between the 4 trained models.
+The survey was conducted through google forms and sent to the students in CSCI 567 (3D User Interfaces) through the piazza message board.
+I took a sample of 5 user maps, and for each one, I produced the ouput for each model and listed those as multiple choice answers.
 
@@
 		\end{tabular}
+
+		\caption{Test images and the outputs from our trained models}
 	\end{table*}
@@
 
+	Expanding this idea to include texture maps could be an interesting direction to go.
+Using this same method, it should be possible to translate a given heightmap into a texturemap.
+This could allow the designer to have control over generated styles like barren desert scenes vs dark horror scenes vs fantastical colorful scenes.
+Gathering training data of texturemaps would likely be the bottleneck.
+
+	Modifying the network to perform real time generation would be an interesting addition.
+This would give the user a more interactive experience in designing the enviroments.
+Current latency with this implementation is not fast enough so new network architectures or approaches would be needed.
+
+	Tunable paramaters would also be an interesting addition.
+Giving the user even more control over the results would lead to a better experience.
+I can imagine parameters like intensity, noise, smoothness could rusult in a lot of usable outcomes.
+
+	Expanding the training set with different means.
+Needing to draw all the usermaps for the training set is time consuming and error/bias prone.
+Coming up with a programmatic way to create the usermap training set could have a large impact on the quality of the results.
+
 	\bibliographystyle{plain}

### 2019-12-10 writing/paper/paper.tex
Write

@@
 	The resulting heightmaps can be imported into popular 3d modeling software including Unity, Unreal, Blender, and World Machine.
-For the purposes of visualizing these results, and because of the limited support for Linux in all of the previously listed softwares, we visualize results using a demo application called Fractscape.
+For the purposes of visualizing these results, and because of the limited support for Linux, we visualize results using a demo application called Fractscape.
 Fractscape includes elevation based texture mapping which happens to work very well for visualizing our terrain elevation data.
-Unity, Unreal, Blender and World Machine have a different workflow for painting textures, so using the generality of fractscape works best for our visualizations \cite{starscenesoftware2006}.
+Unity, Unreal, Blender and World Machine all have different workflows for painting textures, so using the generality of fractscape works best for our visualizations \cite{starscenesoftware2006}.
 

### 2019-12-10 app/server.py
Use CPU for app demo inference

@@ -22,7 +22,8 @@ def generate_heightmap():
     "--name", "two-pix2pix", 
     "--results_dir", "./generate/results", 
     "--model", "pix2pix",
-    "--load_size", "256"])
+    "--load_size", "256",
+    "--gpu_ids", "-1"])
 
   with open("generate/results/two-pix2pix/test_latest/images/A_fake_B.png", "rb") as fh:
       encoded_string = base64.b64encode(fh.read())

### 2019-12-11 writing/paper/paper.tex
Write more and get app running on laptop

@@
 
-	Both Pix2Pix and CycleGAN are conditonal networks based on the deep convolutional generative adversarial network architecture proposed by Radford et al. in 2015 \cite{radford2015unsupervised}.
-
-	The objective function of a conditional generative adversarial networks can be expressed as:
+	The objective function of a standard conditional generative adversarial network can be expressed as:
 
@@
 	Where $G$ is the generator network that is attempting to minimize this objective and D is the discriminator network that tries to maximize it.
+
+	Pix2Pix and CycleGAN both make modifications to this objective function and below we go into those changes.
 
@@
 
+	This cycle consistency loss is then combined with the standard GAN loss for each GAN:
+
 	\begin{align*}
@@
 		&+\ \lambda \mathcal{L}_cyc (G, F)  
+	\end{align*}
+
+	The generators then seek to minimize this objective while the discriminators try to maximize it.
+
+	\begin{align*}
+		G^*, F^* = arg \mathop{min}_{G,F} \mathop{max}_{D_x, D_y} \mathcal{L}(G, F, D_x, D_y)
 	\end{align*}
@@
 	We use an interactive open source map tool that visualizes elevation data gathered from Mapzen's global elevation service \cite{richardson2016mapzen}.
+The data for the map tool is aggrgated from 3 sources, USGS, NASA, and NGA.
+The map is a greyscale map, where the brightness of the pixel corresponds to the relative elevation.
 

### 2019-12-12 writing/paper/paper.tex
Work on website

@@
 \usepackage{floatpag}
+\usepackage{caption}
 
@@
 
+	We recieved 11 responses.
+For all 5 samples, users preferred the output of Pix2Pix over the output of CycleGAN for each one.
+When looking at the results of the original training batch to the augmented training batch, users preferred the ouput of the original batch for 4 samples and the augmented batch for one of the samples.
+
+	\begin{table*}[h]
+		\begin{tabular}{c|c|c|c|c}
+			Sample Number & Original Pix2Pix & Augmented Pix2Pix & Original CycleGAN & Augmented CycleGAN \\
+			\hline
+			1 & 6 & 3 & 1 & 1 \\
+			2 & 8 & 1 & 1 & 1 \\
+			3 & 3 & 6 & 2 & 0 \\
+			4 & 5 & 4 & 1 & 1 \\
+			5 & 7 & 3 & 1 & 0
+		\end{tabular}
+
+	\caption{Survey Results}
+	\end{table*}
+
 	\begin{table*}[t]
@@
 
+	We saw that we can generate good results with generative adversarial networks. It's easy to generate these environments faster and more efficiently than traditional methods.
+
+	Using heightmaps as a target allows for the resulting environments to be generalizable. We can import the heightmaps into many popular 3d software applications like Unity, Unreal and Blender.
+
+	The training time is computationally intensive, but inference is quick.
+Once a model is trained, it's easy to generate as many new environments as you want.
+This has massive improvements over hand crafted environments.
+
+	Pix2Pix performs much better than CycleGAN and users prefer its output for this task over CycleGAN for every sample. Pix2Pix offers specificty that CycleGAN can not match.
+
+	This is a viable area of research, and there's opportunities to tune these models further to get better results.
+Training data is the most important part of this project, and spending more time with it will lead to better results.
+We also show that augmenting the training data is an easy way to expand a dataset, but it can lead to overfitting.
+There is a middleground between too much and too little for augmentation.
+
 	\subsection{Future Work}
@@
 
+	\clearpage
+
+	\begin{figure}[h!]
+	\caption{Sample usermaps, generated heightmaps, and visualized 3d terrain}	
+	\includegraphics[width=2\columnwidth]{viz-1.png} 
+	\end{figure}
+
+	\begin{figure}[h!]
+	\includegraphics[width=2\columnwidth]{viz-2.png} 
+	\end{figure}
+
+	\begin{figure}[h!]
+	\includegraphics[width=2\columnwidth]{viz-3.png} 
+	\end{figure}
+
 \end{document}

### 2019-12-12 writing/paper/paper.tex
Fix typos in paper

@@
 
-	CycleGAN introduces cycle consistency loss to conditional generativeadversarial networks.
+	CycleGAN introduces cycle consistency loss to conditional generative adversarial networks.
 It works by measuring the loss of a mapping from one distribution to another and then back to the original distribution again.
@@
 	Where pix2pix trains one GAN (one generator and one discriminator), CycleGAN trains two, (Generator A and B and Discriminator A and B).
-learning the mapping between the two GANs is the trick to getting unpaired image translation to work.
+Learning the mapping between the two GANs is the trick to getting unpaired image translation to work.
 CycleGAN introduces a cycle consistency loss.
@@
 This leads to an interesting challenge of putting together a realistic yet sizable dataset.
-In order to achieve this, we use terrain data of the earth as opposed to fictional or manufactured terrain data.
+In order to achieve this, we use terrain data of the Earth as opposed to fictional or manufactured terrain data.
 In the next two sections we go into the methodology behind building the dataset.
@@
 	We use an interactive open source map tool that visualizes elevation data gathered from Mapzen's global elevation service \cite{richardson2016mapzen}.
-The data for the map tool is aggrgated from 3 sources, USGS, NASA, and NGA.
+The data for the map tool is aggregated from 3 sources, USGS, NASA, and NGA.
 The map is a greyscale map, where the brightness of the pixel corresponds to the relative elevation.
 
-	We then built a program that controls a headless browser using an open source library developed by Google called puppeteer.
+	We then built a program that controls a headless browser using an open source library developed by Google called Puppeteer.
 This script loads a random coordinate, checks if the coordinate is over land and if so, then we save an image of the elevation data at that coordinate.
@@
 	We then manually build the usermap training set using our user interface application.
-For each image in the heightmap training set, We paint a usermap to represent that heightmap, using the semantically significant brushes that most closely match the terrain (mountains, hills, plains, rivers, and terraces).
+For each image in the heightmap training set, we paint a usermap to represent that heightmap, using the semantically significant brushes that most closely match the terrain (mountains, hills, plains, rivers, and terraces).
 This process is a subjective one and all usermaps were painted by myself.
@@
 	Both models are built on top of the PyTorch framework \cite{paszke2019pytorch} and are trained for 200 epochs.
-We use the open source repo implemented by the Jun-Yan Zhu, who published the research on both models \cite{junyanzhu2017github}.
+We use the default hyperparameters of 0.0002 for the learning rate and ADAM for the optimizer.
+We use the open source repo implemented by Jun-Yan Zhu, who published the research on both models \cite{junyanzhu2017github}.
+Both models are 15 layers deep and contain multiple convolutional layers and residual blocks \cite{isola2016imagetoimage} \cite{zhu2017unpaired}.
 
 	\subsubsection{Pix2Pix}
-		Our pix2pix model contained a total of 57.183 million parameters.
+		Our pix2pix model contains a total of 57.183 million parameters.
 The generator network contains 54.414 million parameters and the discriminator contains 2.769 million parameters.
@@
 
-	Subjective results of networks show that pix2pix performs much better than cyclegan.
+	Subjective results of networks show that Pix2Pix performs much better than CycleGAN.
 For the specfic task of segmented image translation, paired training data allows the network to learn the meaning of the segmentations much better.
@@
 
-	Both models trained on the augmented batch produce much higher fidelity images. Both pix2pix models produce usable and realistic terrain.
+	Both models trained on the augmented batch produce higher fidelity images. Both Pix2Pix models produce usable and realistic terrain.
 
@@
 	We recieved 11 responses.
-For all 5 samples, users preferred the output of Pix2Pix over the output of CycleGAN for each one.
+For all 5 samples, users preferred the output of Pix2Pix over the output of CycleGAN.
 When looking at the results of the original training batch to the augmented training batch, users preferred the ouput of the original batch for 4 samples and the augmented batch for one of the samples.
